---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Read the csv file and then check the structure of the file.
```{r}
activity<-read.csv("./activity/activity.csv", header=TRUE)
str(activity)
```
Reformat the date and check that it is a 'date' class. 
```{r}
library(lubridate)
activity$date<-ymd(activity$date)
str(activity)
```

Re-group the data into dates. 
```{r}
library(dplyr)
tbl_df(activity)
perday<-group_by(activity, date)
```
## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day. 
```{r}
totalperday<-summarise(perday, totalsteps = sum(steps))
library(ggplot2)
qplot(totalsteps, data=totalperday, geom="histogram")
mean(totalperday$totalsteps, na.rm=TRUE)
median(totalperday$totalsteps, na.rm=TRUE)
```

## What is the average daily activity pattern?
Calculate the average number of steps taken per day
```{r}
average<-mutate(group_by(perday, date),averagedsteps=mean(steps, na.rm=TRUE))
```

Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
plot(average$interval, average$averagedsteps, type="l", xlab="interval", ylab="average number of steps taken" )
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)
```{r}
sum(is.na(activity))
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
imputed <- replace(activity, is.na(activity), mean(activity$steps, na.rm = TRUE))
str(imputed)
```

Make a histogram of the total number of steps taken each day and Calculate and report the **mean** and **median** total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
tbl_df(imputed)
totalimputed<- imputed %>% group_by(date) %>% summarise(totalsteps2=sum(steps))
qplot(totalsteps2, data=totalimputed, geom="histogram")
mean(totalimputed$totalsteps2, na.rm=TRUE)
median(totalimputed$totalsteps2, na.rm=TRUE)
```

The median value is different. 
The median takes the mean value.

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
 new<-mutate(totalimputed, day= if (weekdays(totalimputed$date)==TRUE) {c("weekdays")} else {c("weekends")})
```
